Some examples in extended logic counting classes

### 2017-01-12 LICS/appendix-syntactic.tex
Version nueva.

The proof will be separated in two parts. In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$. Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$. 
From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula. In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF. 
We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$. From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.

@@

 Let $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$ and $\bar{X} = (X_1,\ldots,X_{\length{\bar{X}}})$.
To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas. 
Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
$$
\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
$$
Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.

@@

	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
	$$
	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
	$$
	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e. $\varphi_i$ is inconsistent), we do the following. If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
\end{itemize}
It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$. From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c). Furthermore, it has the following form:
$$
\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
$$
where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
\begin{claim}\label{claim:minusone}
For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
\end{claim}
\begin{proof}
Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$. Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$. This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction. 
\end{proof}

The previous claim and proof motivates the following definitions.
For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
$$
\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
$$ 
If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.

Now, we have all the tools to introduce the main formula of the proof. 
Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
\begin{align}
\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
& \hspace{1cm} \Big( \; 
\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\ 
& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
\end{align}
To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$. 
The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).

We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
This formula basically checks if $\varphi_i$ is satisfiable or not.
Then, first define the formula:
$$
\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
$$
This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
$$
\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
$$
This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$. Then we generalize this construction for each $\varphi_i$:
$$
\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
$$
By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.

@@

{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section. Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c). We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.

We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$. If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$. If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined. If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
Now, for each $\varphi_i$ we define:
$$
\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
$$
Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.

>>>>

The proof will be separated in two parts. In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$. Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$. 
From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula. In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF. 
We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$. From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be~proven.

@@

To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in a DNF formula. 
More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i(\bar{X}, \bar{x}) = \exists \bar{y} \, \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ is a conjunction of atomic formulas or negation of atomic formulas. Furthermore, we suppose that each conjunct $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ has the~form:
$$
\varphi_i'(\bar{X}, \bar{x}, \bar{y}) =  \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
$$
Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ will appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.

@@

	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
	$$
	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
	$$
	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  as $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e. $\varphi_i$ is inconsistent), we do the following. If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
\end{itemize}
It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$. From now on, we assume that each $\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \;  \varphi_i'(\bar{X},\bar{x}, \bar{y}) $ satisfies conditions (a), (b) and (c), and $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ has the following~form:
$$
\varphi_i'(\bar{X},\bar{x}, \bar{y}) \; = \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
$$
where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
\begin{claim}\label{claim:minusone}
For a given ordered structure $\A$ and a first-order assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists a second-order assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
\end{claim}
\begin{proof}
Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$, such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$. Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$. This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.

For the other direction, if $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$ for a second order assignment $V$ for $\A$, then it is easy to check $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ since $\varphi^{\fo}_i(\bar{x},\bar{y})$ is a subformula of $\varphi_i'$.
\end{proof}

The previous claim and proof motivates the following definitions.
For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
One can easily check that for every assignments $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all assignments of $\bar{X}$ when $v$ is fixed.
Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
$$
\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
$$ 
If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
Furthermore, let $V^*$ the second order assignment and $v^*$ the first order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.

Now, we have all the tools to introduce one of the main formulas in order to define $\kappa(\alpha)$. 
Intuitively, we want to exclude the assignment $(V^*, v^*)$ from the satisfying assignments of $\varphi_i(\bar{X},\bar{x})$.
Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
$V \neq V^*$ or $v \neq v^*$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
This property can be defined as follows:
\begin{align}
\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
& \hspace{1cm} \Big( \; 
\exists\bar{y}~\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\ 
& \hspace{1.5cm} (\; \exists\bar{x}'\exists\bar{y}\; \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x} \; ) \;\; \Big)  \label{eq:part3}
\end{align}
To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
Then, the conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by defining that if $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ is satisfied then $V^*(\bar{X}) \subsetneq V(\bar{X})$. 
The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
Finally, from the previous discussion one can easily check that $\psi_i(\bar{X},\bar{x})$ satisfies the desire property.

We are ready to define the formula  $\kappa(\alpha)$ as $\sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i^*(\bar{X},\bar{x})$
where each modified disjunct $\varphi_i^*(\bar{X},\bar{x})$ is constructed as follows. 
For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
This formula basically checks if $\varphi_i$ is not satisfiable (recall Claim~\ref{claim:minusone}).
Define the first formula $\varphi_1^*$ as:
$$
\varphi^*_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
$$
This formula accepts all the assignments that satisfy $\varphi_1$, except for the assignment $(V^*,v^*)$ of $\varphi_1$. The second formula $\varphi_2^*$ is defined as:
$$
\varphi^*_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
$$
This models all the assignments that satisfy $\varphi_2$, except for the assignment $(V^*,v^*)$ of $\varphi_1$. Moreover, if $\varphi_1$ is not satisfiable, then $\psi_1(\bar{X},\bar{x})$ and $\chi_1$ will hold, and the formula $\psi_2(\bar{X},\bar{x})$ will forbid the assignment $(V^*,v^*)$ of $\varphi_2$. 
One can easily generalize this construction for each $\varphi_i$ as follows:
$$
\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
$$
From the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.

@@

{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
First, rewrite $\varphi(\bar{X},\bar{x})$ as in the previous section. Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c). Also, consider the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.

We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if, and only if, $\sem{\beta}(\A) = 0$. If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$. If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and define $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$. If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
Now, for each $\varphi_i$ we define:
$$
\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
$$
Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i^*(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
